---
title: "Lab 14 HW"
output: 
  html_document: 
    keep_md: yes
author: "Julie Trundle"
date: '2022-02-28'
---


### Load the Libraries
```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library("palmerpenguins")

#install.packages("ggVennDiagram")
library(ggVennDiagram)
library(RColorBrewer)

#install.packages("ggworldcloud")
library(ggwordcloud)


options(scipen=999) #cancels the use of scientific notation for the session
```


### Data
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))

beachbugs_long <- readr::read_csv("data/beachbugs_long.csv")
```

### 1. 
Clean up the column names (no capitals, not spaces) of `superhero_info`, then use 2 functions to remind yourself of structure of the `superhero_info` data set.

```{r}
superhero_info <- clean_names(superhero_info)
```

```{r}
names(superhero_info)
```

```{r}
glimpse(superhero_info)
```

```{r}
summary(superhero_info)
```

### 2.
Are bad guys bigger? Make box-plots of weight by `alignment` (alignment on the x-axis).
```{r}
superhero_info %>% 
  filter(alignment!="NA") %>% 
  select(weight, alignment) %>% 
  ggplot(aes(x=alignment, y=weight))+
  geom_boxplot()
```
Overall, the distribution of weight does not seem to have a correlation to alignment.  

### 3. 
Now, make a violin plot of weight by `alignment` (alignment on the x-axis). Add some color!
  What information can you observe in the violin plot that was not visible in the boxplot?
```{r}
superhero_info %>% 
  filter(alignment!="NA") %>% 
  select(weight, alignment) %>% 
  ggplot(aes(x=alignment, y=weight, fill=alignment))+
  geom_violin()
```
This shows that there are smaller peaks in the data at some higher weights, but the majority of superheros are all around the same weight and alignment doesn't seem to be correlated to weight. 

### 4. 
Use `alpha = .5` in `geom_boxplot()` and `geom_violin()` to layer both plots on top of one another. What does this tell you about the distribution of weight in "`bad`" guys?
```{r}
superhero_info %>% 
  filter(alignment!="NA") %>% 
  select(weight, alignment) %>% 
  ggplot(aes(x=alignment, y=weight, fill=alignment))+
  geom_violin(alpha=0.5)+
  geom_boxplot(alpha=0.5)
```
Most bad guys have a weight around 100. 

### 5. 
Box plots are great for showing how the distribution of a numeric variable (e.g. weight) varies among a categorical variable (e.g. alignment).
  Make your own violin plot with the superhero data. 
  What is your numeric data? 
  What is your categorical variable?

```{r}
names(superhero_info)
```

```{r}
superhero_info %>% 
  filter(gender!="NA") %>% 
  ggplot(aes(x=gender, y=height, fill=gender))+
  geom_boxplot(alpha=0.5)+
  geom_violin(alpha=0.5)
```

### 6. 
Remind yourself what `beachbugs` looks like. Then generate a heatmap of buglevels by site and year. 
color it with `scale_fill_gradient(low="yellow", high="red")` or colors of your choice. you may find it looks best with one color!
(dont forget, `coord_flip()` is a quick way to improve the look of your plot if you dont like the default orientation)

```{r}
glimpse(beachbugs_long)
```

```{r}
beachbugs_long
```
```{r}
beachbugs_long %>% 
  ggplot(aes(x=site, y=year, fill=buglevels))+
  geom_tile()+
  scale_fill_gradient(low="yellow", high="red")+
  theme(axis.text.x = element_text(angle=60))
```

### 7.  
Use the provided code to normalize the beachbug data set. 
Then make a heatmap with the `beachbugs_normalized` data, and use the same color chois as above. Which heatmap do you think is more informative? why?

```{r}
#makes a new column of the highest buglevel observed at each site
beachbugs_w_max <- beachbugs_long %>% 
  group_by(site) %>% 
  mutate(max_buglevel = max(buglevels, na.rm=T)) %>% 
  arrange(site)
beachbugs_w_max

#makes a new table where the buglevels are normalized to the max_buglevel
beachbugs_normalized <- beachbugs_w_max %>% 
  group_by(site) %>% 
  mutate(norm_buglevel = buglevels/max_buglevel) %>% 
  arrange(site,year) %>%
  select(site, year, norm_buglevel)   # you dont have to select(), but I think its a clearer looking table
beachbugs_normalized
```
 
```{r}
beachbugs_normalized %>% 
  ggplot(aes(x=site, y=year, fill=norm_buglevel))+
  geom_tile()+
  scale_fill_gradient(low="yellow", high="red")+
  theme(axis.text.x = element_text(angle=60))
```
This heat map gives us a wider range of colors, so the difference between buglevels is more obvious than in the first heat map. 

### 8.
Let's make a venn diagram of `superhero_info`, from 4 questions:
Is their alignment evil? 
Are their eyes red?
Are they male?
Are they bald?

Start by making the 4 vectors, then the list of vectors. The vector for alignment is provided:
### super heros venn
```{r}
names(superhero_info)
```

```{r}
 #evil

evil_vec <- superhero_info %>%
  filter(alignment == "bad")%>%
  pull(name)

# red eyes

red_eye_vec <- superhero_info %>% 
  filter(eye_color=="red") %>% 
  pull(name)

# male

male_vec <- superhero_info %>% 
  filter(gender=="male") %>% 
  pull(name)

# bald

bald_vec <- superhero_info %>% 
  filter(hair_color=="bald") %>% 
  pull(name)
```

Your list of vectors will look something like this:
```{r}
questions_list <- list(evil_vec, red_eye_vec, male_vec, bald_vec)
questions_list
```

### 9. 
Let's make the venn diagram! use the code from lab as a reference. 
```{r}
# something like:
 ggVennDiagram(questions_list, category.names = c("evil_vec", "red_eye_vec", "male_vec", "bald_vec"))

```


### 10. Choose one intersection of the venn diagram that is interesting to you. Use dplyr to find the names of the superheros in that intersection. 

```{r}
superhero_info %>% 
  filter(eye_color=="red") %>% 
  filter(alignment=="bad") %>% 
  select(name)
```

### 11. Make another venn diagram with the `superhero_info` data. What are your questions? ( At least 2!)
Do they have blue eyes? 
Are they human?

```{r}
blue_eye_vec <- superhero_info %>% 
  filter(eye_color=="blue") %>% 
  pull(name)

human_vec <- superhero_info %>% 
  filter(race=="human") %>% 
  pull(name)
```


```{r}
blue_human_list <- list(blue_eye_vec, human_vec)
```

```{r}
ggVennDiagram(blue_human_list, category.names = c("blue_eyes", "human"))
```

### 12.
What are some very common super powers? Lets make a word cloud with the `superhero_powers` data.

Use the provided code to make the frequency table, then make a word cloud with it. 
Remember, you can change `scale_size_area(max_size = 20)` to a different number if the words wont fit!

```{r}
head(superhero_powers)
```
```{r}
power_frequency <- superhero_powers %>% 
  select(-hero_names) %>%
  summarise_all(~(sum(.))) %>%
  pivot_longer(everything(), names_to = "power", values_to = "freq")
power_frequency
```


```{r}
power_frequency %>% 
  ggplot(aes(
    label=power,
    size=freq,
    color=power))+
  geom_text_wordcloud()+
  scale_size_area(max_size = 7)+
  theme_minimal()
```
### 13.  
Who are some very powerful supers? 
 Lets make a different word cloud with the `superhero_powers` data. 
Use the provided code to make the frequency table, then make a word cloud with it.
You can use `hero_names` for the labels, and `sum_powers` for size. 

```{r}

power_quantity <- superhero_powers %>% 
  pivot_longer(-hero_names, names_to = "power", values_to = "yes_or_no") %>% 
  group_by(hero_names) %>% 
  mutate(sum_powers = sum(yes_or_no, na.rm=T)) %>%
  arrange(desc(sum_powers), hero_names, desc(yes_or_no))
power_quantity

#we have to trim down to only the top 50, or it will crowd the image!
power_quantity <- power_quantity %>% 
  ungroup %>%
  distinct(hero_names, sum_powers) %>%
  slice_max(sum_powers, n = 50)
power_quantity

```



```{r}
power_quantity %>% 
  ggplot(aes(
    label=hero_names,
    size=sum_powers,
    color=hero_names
  ))+
  geom_text_wordcloud()+
  scale_size_area(max_size = 5)
```

## That's it! 🎉
Thanks for coding with us all winter! 
Now go finish your group project :) 

